# Reproducible Research - Peer-graded Assignment: Course Project 1


## Loading and preprocessing the data

1. Load the data (i.e. read.csv())

```{r setoptions, echo=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/Tim/Documents/Data Science/data")
knitr::knit_hooks$set(inline = function(x) {if(!is.numeric(x)){ x } else{prettyNum(round(x, 0), big.mark=",")}}) 
```

```{r libraries, echo=FALSE, results='hide'}
library(ggplot2)
library(scales)
library(Hmisc)
```


```{r dataLoad}
if(!file.exists('activity.csv')){
      unzip('repdata_data_activity.zip')
}
activity <- read.csv('activity.csv')
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r dateFormat}
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r stepsPerDay}
stepsPerDay <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
```

2. Make a histogram of the total number of steps taken each day

```{r Hist1}
qplot(stepsPerDay, xlab = "Steps per Day", ylab = "No. of Days", binwidth = 400)
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r mean&median1}
stepsPerDayMean <- mean(stepsPerDay)  
stepsPerDayMedian <- median(stepsPerDay)
```

Mean steps per day: `r stepsPerDayMean`  
Median steps per day: `r stepsPerDayMedian`


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r timeSeries1}
avgSteps <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
plot(avgSteps$interval, avgSteps$steps, type = "l", lwd = 2, col = "skyblue", 
     main = "Time series of 5-minute interval vs. steps taken", xlab = "interval", 
     ylab = "avg steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r maxSteps}
maxInt <- avgSteps$interval[which.max(avgSteps$steps)]
```

Interval containing maximum number of steps: `r maxInt`


# Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r NAs}
NAs <- sum(is.na(activity))
```

Total number of rows with NAs: `r NAs`

2. Devise a strategy for filling in all of the missing values in the dataset.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r NAsFilled}
activityImputed <- activity
activityImputed$steps <- impute(activity$steps, fun = mean)
```

4. Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.

```{r Hist2}
stepsPerDayImputed <- tapply(activityImputed$steps, activityImputed$date, FUN = sum)
qplot(stepsPerDayImputed, xlab = "Steps per Day (Imputed)", ylab = "No. of Days", binwidth = 400)
```

```{r mean&median2}
stepsPerDayImputedMean <- mean(stepsPerDayImputed)
stepsPerDayImputedMedian <- median(stepsPerDayImputed)
```

Mean imputed steps per day: `r stepsPerDayImputedMean`  
Median imputed steps per day: `r stepsPerDayImputedMedian`


# Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r dayType}
activityImputed$day <- weekdays(activityImputed$date)
activityImputed$dayType <- "weekday"
activityImputed[activityImputed$day == "Saturday" | activityImputed$day == "Sunday", ]$dayType <- "weekend"
activityImputed$dayType <- factor(activityImputed$dayType)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r timeSeries2}
avgStepsImputed <- aggregate(steps ~ interval + dayType, data = activityImputed, FUN = mean)
ggplot(avgStepsImputed, aes(interval, steps)) +
      geom_line(color = "springgreen") +
      facet_grid(dayType ~ .) +
      xlab("interval") +
      ylab("avg steps") +
      theme(panel.background = element_rect(fill = "lightcyan"))
```